HTW Model

### 2024-02-20 Model/htw_model.qmd
writing progress; remove files from failed render

The residuals of the model predictions for the testing stage (@fig-htw-resid-pred) also show a predictable pattern of fitting method - with models fit only to the test data showing the best performance, followed by models fit to both training and test data, and with models fit only to the training data showing the worst performance (note that y axes are scaled different between plots). Unsurprisingly, the advantage of EXAM is strongest for extrapolation positions (the three smallest bands for both groups - as well as the two highest bands for the Constant group). Although EXAM tends to perform better for both Constant and Varied participants (see also @tbl-htw-modelError), the relative advantage of EXAM is generally larger for the Constant group - a pattern consistent across all three fitting methods. Panel B of @fig-htw-resid-pred directly compares the aggregated observed data to the posterior predictive distributions for the testing stage. Of interest are a) the extent to which the median estimates of the ALM and EXAM posteriors deviate from the observed medians for each velocity band; b) the ability of ALM and EXAM to discriminate between velocity bands; c) the relative performance of models that are constrained by the training data (i.e. the 'fit to train only' and 'fit to both' models) compared to the 'fit to test only' models; and d) the extent to which the variance of the posterior predictive distributions mimics the variance of the observed data.

>>>>

The residuals of the model predictions for the testing stage (@fig-htw-resid-pred) also show a sensible pattern across fitting methods - with models fit only to the test data showing the best performance, followed by models fit to both training and test data, and with models fit only to the training data showing the worst performance (note that y axes are scaled different between plots). Unsurprisingly, the advantage of EXAM is strongest for extrapolation positions (the three smallest bands for both groups - as well as the two highest bands for the Constant group). Although EXAM tends to perform better for both Constant and Varied participants (see also @tbl-htw-modelError), the relative advantage of EXAM is generally larger for the Constant group - a pattern consistent across all three fitting methods.

Panel B of @fig-htw-resid-pred directly compares the aggregated observed data to the posterior predictive distributions for the testing stage. Of interest are a) the extent to which the median estimates of the ALM and EXAM posteriors deviate from the observed medians for each velocity band; b) the ability of ALM and EXAM to discriminate between velocity bands; c) the relative performance of models that are constrained by the training data (i.e. the 'fit to train only' and 'fit to both' models) compared to the 'fit to test only' models; and d) the extent to which the variance of the posterior predictive distributions mimics the variance of the observed data.

Considering first the models fit to only the testing data, which reflect the best possible performance of ALM and EXAM at capturing the group-aggregated testing patterns. For the varied group, both ALM and EXAM are able to capture the median values of the observed data within the 66% credible intervals, and the spread of model predictions generally matches that of the observed data. For the constant group, only EXAM is able to capture the median range of values across the velocity bands, with ALM generally underestimating human velocoties in the upper bands, and overestimating in the lower bands. In the case of band 100, the median ALM prediction appears to match that of our participants - however this is due to a large subset of participants have ALM predictions near 0 for band 100, a pattern we will explore further in our considertation of individual patterns below.
